---
title: "treeTests_troubleshoot_subsetting"
author: "Janet Young\n"
date: "`r Sys.Date()`\n"
output: github_document
always_allow_html: true
---

```{r setup, include=TRUE}

library(ape)
library(ggtree)
library(tidytree)

### this is to let the doc show the errors and continue to knit
knitr::opts_chunk$set(echo = TRUE, error = TRUE)
# override error output hook here
knitr::knit_hooks$set(error = function(x, options) { 
  knitr::knit_exit()
})
```

# Get example tree data

Make a tree

```{r}
tr <- rtree(10)
```

Make table of information about tips

```{r}
tree_info <- data.frame(label = paste0("t",1:10),
                        group_character = c(rep("group1", 3),
                                               rep("group2", 4),
                                               rep("group3", 3)))

## add factor column:
tree_info[,"group_factor"] <- factor(tree_info[,"group_character"])

## add integer column:
tree_info[,"group_integer"] <- as.integer(gsub("group","",tree_info[,"group_character"]))

## add TRUE/FALSE column:
tree_info[,"in_group2"] <- tree_info[,"group_character"] == "group2"

tree_info
```

Join info and tree together:

```{r}
tr_plus_info <- left_join(tr, tree_info, by="label")
```

# Now plot

Plot, using `in_group2` column (TRUE/FALSE) to subset geom_tippoint - works:

```{r, fig.height=4, fig.width=4}
ggtree(tr_plus_info) +
    geom_tiplab(aes(color=group_character), offset=0.05) +
    geom_tippoint(aes(subset=in_group2, 
                      color=group_character),
                  size=3, 
                  show.legend = FALSE)
```



Plot, using a logical test on the `group_integer` column (group_integer==2) to subset geom_tippoint - this works:

```{r}
ggtree(tr_plus_info) +
    geom_tiplab(aes(color=group_character), offset=0.05) +
    geom_tippoint(aes(subset= (group_integer==2), 
                      color=group_character),
                  size=3, 
                  show.legend = FALSE)
```


Plot, using a logical test on the `group_factor` column (group_factor=="group2") to subset geom_tippoint - this fails:


```{r}
ggtree(tr_plus_info) +
    geom_tiplab(aes(color=group_character), offset=0.05) +
    geom_tippoint(aes(subset= (group_factor=="group2"), 
                      color=group_character),
                  size=3, 
                  show.legend = FALSE)
```

Plot, using a logical test on the `group_character` column (group_character=="group2") to subset geom_tippoint - this fails:


```{r, fig.height=4, fig.width=4}
ggtree(tr_plus_info) +
    geom_tiplab(aes(subset= (group_character=="group2"),
                    color=group_character), offset=0.05) +
    geom_tippoint(aes(subset=in_group2, 
                      color=group_character),
                  size=3, 
                  show.legend = FALSE)
```

# Finished

```{r}
sessionInfo()
```

